---
title: "Association between Crop burning and Air quality in South Asian Nations"
author: "Sujan Bhattarai"
format:
  html:
    code-fold: true
    code-summary: "Show the code"
---

 <style>
        img {
            max-width: 70%;
            height: auto;
            display: block;
            margin-left: auto;
            margin-right: auto;
            float: left;
        }
        
        figcaption {
            margin-top: 10px;
            text-align: center;
        }
 </style>

```{r setup, include=FALSE}
#| code-fold: true
knitr::opts_chunk$set(
  comment = "#>",
  warning = FALSE,
  message = FALSE
)
```

Every winter, South Asian cities goes through extreme air pollution during winter. There could be several reason for it, but one major reason is due to crop burning in the agricultural land. The other reason is: during winter, there is low circulation of wind resulted by fog, causing the city dusts to not escape.
As a part of this project, I will check the strength of the relationships between
crop burning with air quality index value in South Asian Nation. 

As a part of this analysis, I will investigate two major questions:
 - study the relationships between crop burning and air quality index, generalized for all south Asia
 - identify south asian country with highest crop burning and its contribution to air quality
 - use forcasting to predict where the country is heading too in coming years
 
#### Setup
All the required libraries(based on previous knowledge) is loaded in the first chunk. Some libraries will be loaded in their
individual chunks to show what function is being used from what package.
```{r}
#| code-fold: true
##Load all required libraries
library(tidyverse)
library(readr)
library(gt)
library(tufte)
library(feasts)
library(here)
library(janitor)
```


## Data collection and wrangling
The required data are downloaded from World bank data catalogue, Kaggle website, and ClimateWatch. The dataset are cleaned based on requirement using stringR package, dplyr, and tidyverse. In the first part, the emission data from the kaggle is cleaned and joined with air quality data from worldbank. The data are summarized in yearly basis, ranging from 1990 to 2020.
The detailed data analysis is shown in the code chunks.

```{r}
##read the data and clean the name 
data_world_bank <- read_csv(here("posts", "southAsia_crop","data",
                                 "Agrofood_co2_emission.csv"),
                                 show_col_types = FALSE) %>%
                   janitor::clean_names()

#filter data for south asia
south_asian_countries <- c("Nepal", "India", "Bangladesh", "Sri Lanka", "Pakistan",
                          "Afghanistan", "Maldives")
                          
#just filter for south asian region
south_asian_data <- data_world_bank %>% 
                    filter(area %in% c(south_asian_countries)) %>% 
                    mutate(year = as.numeric(year))

```


```{r}
##read the data and transpose it to form a database
#link to the data: 
aqi_world_bank1 <- read_csv(here("posts", "southAsia_crop","data", "aqi.csv")) %>% 
                   clean_names() %>% 
                   t()

##Perform data cleaning 
aqi_world_bank <- data.frame(cbind(year1 = rownames(aqi_world_bank1), 
                                   aqi_world_bank1)) %>%  
                  mutate(year = as.numeric(sub(".*yr(\\d{4}).*", "\\1", year1)))

##filter only those rows where column is not equal to NA

aqi_world_bank <- aqi_world_bank %>% 
                  filter(!is.na(year))

##select only required columns and rename bad-columns 
aqi_world_bank <- aqi_world_bank %>% 
                  select(year = year, 'aqi_value' = 'V2') %>% 
                  rownames_to_column(var = "RowID") %>% #drop the index, not required
                  select(-RowID)

##arrange the data based on numeric columns
aqi_world_bank <- aqi_world_bank %>% 
                  arrange(year) %>% 
                  mutate(year = as.numeric(year))
```


```{r}
##filter columns from the first dataset 
SA_reduced_columns <- south_asian_data %>% 
                      select(area, year, crop_residues, average_temperature_c)

##group by year and summarize for all countries
SA_summarized  <- SA_reduced_columns %>% 
                      group_by(year) %>% 
                      summarize(crop_residues = sum(crop_residues),
                                temp_average  = mean(average_temperature_c))

##left join with aqi world bank 
SA_emission <- left_join(SA_summarized, aqi_world_bank, by = "year") %>% 
               mutate(aqi_value = as.numeric(aqi_value))

```

## Data Exploration
The cleaned data is visualized using graphs. This is an exploratory phase to just observe the general trend and types of data
in the dataframe. This is useful in later analysis. Data exploration is important to understand missingness and errors in
data. The graph below is the long-term trend graph of crop emission in south asian countries. The graph appears to show strong increasing trend in crop burning.

```{r}
# perform exploratory data analysis 
ggplot(SA_emission, aes(x = year, crop_residues)) +
  geom_line (stat = "identity", fill = "skyblue", color = "black") +
  labs(title = "Emission from Crop Residues burning in South Asian Countries",
       x = "Year",
       y = "Emission (Metric ton)") +
  theme_minimal()

```
Similarly, this graph is the trend for air quality value in south asian nation. When the x-axis scale is at 1 unit interval, 
the graph looks much like a white noise, but it can have trend. This is where moving average becomes useful. I used 10years moving average to reconstruct this graph and it shows clear increasing trend.

```{r}
# include temperature with aqi index, adding interaction between crop residues
ggplot(SA_emission, aes(year, aqi_value)) +
  geom_line(color = "blue") +
  labs(title = "Average air quality increase in South Asian Countries since 1990",
       x = "Year",
       y = "AQI value") +
  ylim(c(0, 10))+
  theme_minimal()

```

## Data Transformation
The data exploration offered insights on this data. The next step involves transformation. In this phase, data will be
transformed mathematically so they satisfies the law of the model I intend to use. Since, I would be fitting linear regression model, its important to have association of x with y. So, i transformed my predictor variable(crop emission) using rolling average so that there exists strong correlation.

```{r,   collapse = TRUE}
#correlation between these two plots, when no transformation
cor(SA_emission$aqi_value, SA_emission$crop_residues)
```

```{r,   collapse = TRUE}
#Try transforming data, or smoother the data to detect the pattern
library(slider)
SA_emission <- SA_emission %>% 
               mutate(aqi_avearge = slide_dbl(aqi_value, mean, .before = 10, 
                                              .complete = TRUE))
cor(SA_emission$aqi_avearge, SA_emission$crop_residues, use = "na.or.complete")
```
## Fitting a Linear Model

The data is ready, I propose a statistical hypothesis for my linear regression model:
-   null hypothesis: no change in aqi value with crop residues
-   alternate hypothesis: change in aqi value with crop residues

```{r}
#| code-fold: false
MODEL = lm(data = SA_emission, aqi_avearge ~ crop_residues)
summary(MODEL)
```

Based on the model output, there is statistically significant relationships between air quality index and crop residues, showing the change occured with changed AQI value. This was expected since air quality is directly influenced by smokes and dust from crop burning. The next part involves exploring the country with highest emission.

## Part B
The next steps involves finding the south Asian country has the highest emission. For this, I followed similar steps as for the first part. However, there is small change in data transformation method. Instead of using rolling mean, I normalized data by total cultivable area. To be more precise, the graph below is emission based on country. However, this is biased becuase India has greater area and greater population, so the net total emission will be obviously larger. In that case, if the emission value is divided by total cultivable area within that county, I get emission per unit cultivable area. This is more precise and gives better answer.

```{r}
#further data exploration
ggplot(south_asian_data, aes(year, crop_residues, color = area)) +
  geom_point() +
  facet_grid(cols = vars(area)) +
  labs(
    x = "Year 1990-2020",
    y = "Emission from Crop Burning"
  ) +
  theme(
      # Remove x-axis title
    axis.text.x = element_blank()    # Remove x-axis labels
  )+
  scale_x_continuous(breaks = seq(min(south_asian_data$year), max(south_asian_data$year), by = 30))  # Adjust the breaks as needed

# Save the plot
ggsave("explore.png", width = 10, height = 5)
```

So, if I normalize the emission, I get the following graph. Result: it's not India with highest emission but Bangladesh. 

```{r, collapse=TRUE}
#create a dataframe including values for agricultural land in these countries
total_agri_area <- read_csv(here("posts", "southAsia_crop","data", "area_sa.csv"))

colnames(total_agri_area)

total_agri_area <- total_agri_area %>%  select(-c(`Country Code`, `Indicator Code`)) %>%                    filter(`Country Name` %in% south_asian_countries)

##Pivot longer the data
transposed <- total_agri_area %>% pivot_longer(cols = "1960": "2022") %>% 
              na.omit() %>% 
              filter(name  >= 1990) %>% 
              rename("area_cultivable"= "value", "year"= "name", 'area' = 'Country Name') %>% 
  mutate(year = as.numeric(year))
transposed
                   
```

```{r}
# Divide crop residues by total agricultural land 
##SA_asian_data has the required ID for the division
normalized_area <- left_join(south_asian_data, transposed, by = c("area", "year")) 

normalized_area <- normalized_area %>% 
                   mutate( norm_crop_residue = crop_residues/area_cultivable * 10^6)
```

```{r}
## now normalize the area
ggplot(normalized_area, aes(year, norm_crop_residue, color = area))+
  geom_point()+
  facet_grid(cols = vars(area), scales = "free_x")+
  theme_bw()+
  theme(axis.text.x = element_blank())+
   scale_color_discrete(name = "Country") + 
 labs(
    x = "Year 1990-2020",
    y = "Emission from Crop Burning per cultivable area"
  ) +
  theme(
      # Remove x-axis title
    axis.text.x = element_blank()    # Remove x-axis labels
  )+
  scale_x_continuous(breaks = seq(min(south_asian_data$year), max(south_asian_data$year), by = 30))  # Adjust the breaks as needed

# Save the plot
ggsave("explore2.png", width = 10, height = 5)
```

The next part of my analysis now focus on Bangladesh because it's a country with highest emission per cultivable land. I fit the linear regression model. Bangladesh has the greatest emissions from crop residues in South Asia. So, how strong is the relationship between crop residues and air quality?

```{r}
#  work for bangladesh
data <- c(52.90993831, 51.64362616,
53.74695817, 55.06194439, 56.41649854, 60.5358739, 68.94271136, 66.42738113, 73.27082296, 68.97033719, 67.7673598, 62.75318316, 63.27546257)

# Duplicate the first two values four times each
data <- c(rep(data[1:5], each = 4), data[3:length(data)])

# Create a data frame with a single column
Bangladesh <- data.frame(aqi_value = data)
```

```{r}
#combine this with SA Data for Bangladesh only
bangladesh_filtered <- normalized_area  %>%
                       filter(area ==  "Bangladesh") %>% 
                       cbind(Bangladesh)
```

```{r}
##test the model here
#| code-fold: false
model_bangladesh <- lm(data = bangladesh_filtered, aqi_value ~ crop_residues)
summary(model_bangladesh)
```
Based on the model, for every 0.008 ton increase in crop residues, Air quality Index value increase by 1 unit. The model Residual square value is around 62%. The p values of less than 5% and residual squared error value of 61% provides clue that the model performed well with the variables.

## Further Exploration
Upto this point, I used annual average data. Now, I consider using monthly time series to uncover the trend of air quality in Bangladesh. The process involves decomposing time series and using couple of forecasting model to examine the best model for the dataset.

```{r}
##read all year data and combine them using list methods
files_list <- list.files(here("posts", "southAsia_crop", "data"), 
                         pattern = "^Dhaka.*\\.csv$")
daily_aqi <- data.frame()

# Loop through each file name and read the data
for (file in files_list) {
  file_path <- here::here("posts", "southAsia_crop", "data", file)  # Adjust the path accordingly
  x <- readr::read_csv(file_path)
  daily_aqi <- dplyr::bind_rows(daily_aqi, x)
}
  
##filter where all air quality index values are -999
daily_aqi_filtered <- daily_aqi  %>% 
                      filter(!AQI == -999)
```

It's not useful to conduct time-series analysis by a Daily basis, because crop burning depends every month and not daily. For example: villagers might burn crop residues only in those where there's less wind. And since thousands of villagers exist, they will burn it on a different day. But, there is no online source where I could find monthly data of crop emission. So, we will plot time series for montly air quality.

```{r message = FALSE}
##load feasts package, these contains mostly time series functions 
library(feasts)
library(tsibble)
library(forecast)

##change date format
daily_aqi_filtered <- daily_aqi_filtered %>%
                      mutate(date_column = as.Date(daily_aqi_filtered$`Date (LT)`, 
                            format = "%m/%d/%Y %H:%M")) %>% 
                      na.omit()

plot(daily_aqi_filtered$AQI)
```

```{r}
# Aggregate by month
monthly_aggregated <- daily_aqi_filtered %>%
                      mutate(month = month(date_column),
                             year = year(date_column)) %>% 
                      group_by(year, month) %>% 
                      summarize(aqi_sum = mean(AQI)) %>% 
                      mutate(date_column = make_date(year, month))

##convert this to tsibble
df <- monthly_aggregated %>%
      mutate(date = yearmonth(date_column))
df <- as_tsibble(df, index = date)
df <- df[c('date', 'aqi_sum')]

df %>%
  stl(t.window = 13, s.window = "periodic", robust = TRUE) %>%
  autoplot() +
  ggtitle("STL Decomposition") +
  labs(x = "Time", y = "Value") +
  theme_minimal() +
  theme(plot.title = element_text(size = 16, face = "bold"),
        axis.title = element_text(size = 14),
        axis.text = element_text(size = 12),
        legend.title = element_blank(),
        legend.text = element_text(size = 12))
```

For this time series, I didn't expect seasonality based on my knowledge. But, after decomposing and observing the graph, I see that I was wrong. There's indeed a seasonality, though its contribution to the data is pretty small(this information is given by the bar that is at right of each of the graph). 

## Forecasting
The given R codes in chunks below are for predicting air quality index in Bangladesh over the next five years using four different models: Exponential Smoothing State Space Models (ETS), AutoRegressive Integrated Moving Average (ARIMA), Nonlinear AutoRegressive Neural Network (NNAR), and Trigonometric seasonality, Box-Cox transformation, ARMA errors, Trend and Seasonal components (TBATS). The data for training goes up to the year 2010, and forecasts are made for 20 time points (5 years). The individual predictions from each model are combined by averaging them, creating a combined forecast. This combined forecast, along with the original data and predictions from each individual model, is then presented graphically using the autoplot function. The aim is to evaluate how well these models perform and if combining them improves the accuracy of predictions.

```{r}
##preapre train model and fit the forecasting model
Bangladesh_crop <- ts(bangladesh_filtered$crop_residues, start=1990, frequency=1)

train <- window(Bangladesh_crop, end=2010)
h     <- 20  # Forecasting for the next 5 years (12 months per year)
ETS   <- forecast(ets(train), h=h)
ARIMA <- forecast(auto.arima(train, lambda=0, biasadj=TRUE), h=h)
NNAR  <- forecast(nnetar(train), h=h)
TBATS <- forecast(tbats(train, biasadj=TRUE), h=h)

# Combine forecasts for the next 5 years
Combination <- (ETS[["mean"]] + ARIMA[["mean"]] +
                NNAR[["mean"]] + TBATS[["mean"]]) / 4 
```

```{r}
autoplot(Bangladesh_crop) +
  autolayer(ETS, series="ETS", PI=FALSE) +
  autolayer(ARIMA, series="ARIMA", PI=FALSE) +
  autolayer(NNAR, series="NNAR", PI=FALSE) +
  autolayer(TBATS, series="TBATS", PI=FALSE) +
  autolayer(Combination, series="Combination") +
  xlab("Year") + ylab("Emission from Crop Residues") +
  ggtitle("Forecast crop emission in Bangladesh")+
  theme_bw()
```
Based on the model, the yellow line fits appears closest the the real trend. This is a combination model and it's prediction can be used for the year 2030. Based on this output, Bangladesh will have increased emission above 4000 ton by 2030.

## Next steps
- assessing the accuracy of each of the model
- collect monthly data for crop emission to see if winter contributes further to AQI in south asia


